Estimating the contribution of selection to temporal allele frequency change

### 2019-09-21 manuscript/manuscript.tex
with gc changes
updates to ms

  Adaptation can proceed in natural populations and selection experiments over
  very short timescales.  However, detecting these changes genome-wide is
  difficult, as rapid adaptation often proceeds from standing variation or is
  the result of polygenic selection, both of which may leave faint genomic
  signals indistinguishable from a noisy background of genetic drift. One
  promising signal comes from the genome-wide covariance between allele
  frequency changes observable from temporal genomic data, e.g.
  evolve-and-resequence studies. These temporal covariances reflect how the
  change in neutral allele frequency at one timepoint is predictive of the
  changes at later timepoints when there is heritable fitness variation in the
  population, as neutral alleles can remain associated with their random 
  selected background over time. Since genetic drift does not lead to temporal
  covariance, we can use these covariances to begin to directly estimate what
  fraction of the variation in allele frequency change through time is drift by
  linked selection. Here, we reanalyze two \emph{Drosophila simulans}
  evolve-and-resequence studies, and one artificial selection experiment in
  mice, to quantify the effects of linked selection over short timescales.
  Additionally, taking advantage of selection on replicated populations, we can
  directly estimate the covariance in allele frequency between replicates due
  to convergent selection pressure.  We also identify signals of negative
  temporal covariance that show that selection reverses its direction for a
  reasonable proportion of loci over the time course of a selection experiment.
  Overall, we find that in the three studies we analyzed, linked selection has
  a remarkable impact on short term allele frequency dynamics that previously
  may have been attributed to genetic drift

@@

A long-standing problem in evolutionary genetics is quantifying the roles
genetic drift and selection have in shaping genome-wide allele frequency
changes. Selection can both directly and indirectly affect allele frequency
change, with the indirect effect coming from the action of selection on
correlated loci elsewhere in genome e.g. linked selection
(\cite{Maynard_Smith1974-lc}, \cite{Charlesworth1993-gb,Nordborg1996-nq}; see
\cite{Barton2000-zg} for a review).  Previous work on this question has mostly
focused on teasing apart the impacts of drift and selection on genome-wide
diversity using population samples from a single contemporary timepoint, often
by modeling the correlation between regional recombination rate, gene density,
and diversity created in the presence of linked selection
\parencite{Begun1992-ey,Elyashiv2016-vt}. This approach has shown linked
selection has a major role in shaping patterns of genome-wide diversity across
the genomes of some of organisms
\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\cite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt}.
However, we lack an understanding of how genome-wide linked selection acts over
an important dimension: time.

Presently, there are numerous examples of rapid phenotypic adaptation
\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid
genomic evolution in asexual populations
\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
of fitness makes detecting the impact selection has on genome-wide variation
over short timescales in sexual populations remarkably difficult. This is
because the effect of selection on a polygenic trait (such as fitness) is
distributed across loci proportional to their effect sizes, which can lead to
subtle allele frequency shifts on standing variation that are difficult to
distinguish from background levels of genetic drift and sampling variance.
However, increasingly genomic experimental evolution studies with multiple
timepoints, and in some cases multiple replicate populations, are being used to
detect selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate
modes of selection \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.
In addition these temporal-genomic studies have begun in wild populations, some
with the goal of finding variants that exhibit frequency changes consistent
with fluctuating selection \parencite{Bergland2014-ij,Machado2018-cs}. \vb{In a
  previous paper, we proposed that one useful signal for understanding the
  genome-wide impact of polygenic linked selection detectable from temporal
genomic studies} is the temporal autocovariance in allele frequency changes
\parencite{Buffalo2019-io}.  These covariances are directly estimable from
temporal genomic data and are created when the loci that underly heritable
fitness variation perturb linked neutral alleles; in contrast, when genetic
drift acts alone in a closed population, these covariances are zero in
expectation. Mathematically, temporal covariances are useful because it is
natural to decompose the total variance in allele frequency change across a set
of time intervals into the variances and covariances in allele frequency change
among time intervals.  Furthermore, biologically, these covariances reflect the
extent to which allele frequency changes in one generation predict changes in
another due to a shared fitness background and shared selection pressures.

Here, we provide the first empirical analyses to quantify the impact of linked
selection acting over short timescales (tens of generations) across two evolve
and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and one
artificial selection experiment \parencite{Castro2019-uk}. We repeatedly find a
signal of temporal covariance, consistent with linked selection acting to
significantly perturb genome-wide allele frequency changes across the genome in
a manner that other approaches would not be able differentiate from genetic
drift. We estimate the lower bound of the proportion of total variation in
allele frequency change caused by linked selection, and the correlation between
allele frequency changes between replicate populations caused by response to
convergent selection pressures. Overall, we demonstrate that linked selection
has a powerful role in shaping genome-wide allele frequency changes over very
short timescales.

@@

\begin{center}
  \begin{tabular}{l c c c c c c}
    Study & Species & Selection & Replicates & Pop. Size & Gens. & Timepoints \\ 
  \hline
    \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab & 3 & $\sim$1100 & 14 & 2 \\ \hline
    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab & 10 & $\sim$1000 & 60 & 7 \\ \hline
    \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 20 & 2 \\
 \hline
\end{tabular}
\end{center}

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature lab environment and
evolved for 60 generations, and sequenced every ten generations. Using the
seven timepoints and ten replicate populations, we estimated the genome-wide $6
\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
replicates. Each row of these matrices represent the temporal covariance
$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
change (in ten generation intervals, denoted $\Delta_{_{10}} p_t$) in some
initial reference generation $s$ (the row of the matrix), and some later
timepoint $t$ (the column of the matrix). We corrected these matrices for
biases created due to sampling noise, and normalize the entries for
heterozygosity (see Supplementary Material Sections
\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction} for details on
the bias correction). The covariances are expected to be zero when only drift
is acting in a closed population, as only heritable variation for fitness can
create covariance between allele frequency changes \parencite{Buffalo2019-io}.
Averaging across the ten replicate temporal covariances matrices, we find
temporal covariances that are statistically significant (95\% block bootstraps
CIs do not contain zero), consistent with linked selection perturbing
genome-wide allele frequency changes over very short time periods. The
covariances between all adjacent time intervals are positive and then decay
towards zero as we look at more distant time intervals, as expected when
directional selection affects linked variants' frequency trajectories until
ultimately linkage disequilibrium and the additive genetic variance for fitness
associated with neutral alleles decays \parencite{Buffalo2019-io}. The temporal
covariances per replicate are noisier but this general pattern holds; see
Supplementary Figures \ref{suppfig:barghi-cov-panels}.
\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
time intervals are on average ten generations apart, and given the temporal
decay in covariance we see, the covariances on shorter time-scales (e.g. if
adjacent generations had been sequenced) may well be higher yet (see
Supplementary Material Section \ref{supp:barghi-covs} for more details).

@@

While the presence of positive temporal covariances is consistent with linked
selection affecting allele frequencies over time, this measure not easily
interpretable. Additionally, we can quantify the impact of linked selection on
allele frequency change as the ratio of total covariance in allele frequency
change to the total variance in allele frequency change. Since the total
variation in allele frequency change can be partitioned into variance and
covariance components, $\var(p_t - p_0) = \sum_{i \ne j} \cov(\Delta p_i,
\Delta p_j) / \var(p_t - p_0)$, and the covariances are zero when drift acts
alone, this is a lower bound on how much of the variance in allele frequency
change is caused by linked selection \parencite{Buffalo2019-io}. We call this
measure $G(t)$, which is this measure of the effect of linked selection
between the initial generation $0$ and some later generation $t$, which can be
varied to see how this quantity grows through time. As with the temporal
covariances, the study design of \textcite{Barghi2019-qy} leads our measure
$G(t)$ to be even more conservative, since the temporal covariances with each
ten-generation block between sequenced timepoints are not directly observable,
and are not included in the numerator of $G(t)$. Still, we find a remarkably
strong signal that greater than $20\%$ of total variation in allele frequency
change over 60 generations is directly the result of linked selection.

The replicate design of \textcite{Barghi2019-qy} also allows us to quantify
another covariance: the covariance in allele frequency change between replicate
populations experiencing convergent selection pressures. These
between-replicate covariances are created in the same way as temporal
covariances: neutral alleles linked to a particular fitness background
experience are expected to have allele frequency changes in the same direction
if the selection pressures are similar. Intuitively, where temporal covariances
reflect that neutral alleles associated with heritable fitness backgrounds are
predictive of frequency changes between generations, replicate covariances
reflect that heritable fitness backgrounds common to each replicate predict
(under the same selection pressures) frequency changes between replicates. We
measure this through a statistic similar to a correlation, which we call the
convergent correlation: the ratio of average between-replicate covariance
across all pairs to the average standard deviation across all pairs of
replicates,

@@

A benefit of between-replicate covariances is unlike temporal covariances,
these can be calculated with only two sequenced timepoints and a replicated
study design. This allowed us to assess the impact of linked selection in
driving convergent patterns of allele frequency change across replicate
populations in two other studies. First, we reanalyzed the selection experiment
of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
environment. Since each replicate was exposed to the same selection pressure
and share linkage disequilibria common to the original natural founding
population, we expected each of the three replicate populations to have
positive between-replicate covariances. We find all three pairwise
between-replicate covariances are positive and statistical significant (95\%
CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
convergent correlation coefficient across these replicates as 0.36 ($95\%$
block-bootstrap confidence interval $[0.31, 0.40]$).

%\gc{TODO GIVE conversion \% var due to linked selection. }
Second, we reanalyzed the Longshanks selection experiment, which selected for
longer tibiae length relative to body size in mice, leading to a response of
selection of about 5 standard deviations over the course of twenty generations
\parencite{Castro2019-uk}. This study includes two independent selection lines,
Longshanks 1 and 2 (LS1 and LS2), and an unselected control line (Ctrl).
Consequently, this selection experiment offers a useful control to test our
between-replicate covariances: we expect to see positive between-replicate
covariance in the comparison between the two Longshanks selection lines, but
not between the two pairwise comparisons between the control line and the two
Longshanks lines. We find that this is the case (gray confidence intervals in
Figure \ref{fig:figure-2} C), with the two Longshanks comparisons to the
control line not being significantly different from zero, while the comparison
between the two Longshanks line is statistically significantly different from
zero (CIs [0.013, 0.037]).

@@

\begin{figure}
  \begin{center}
  \includegraphics[]{figures/figure-3.pdf}

  \caption{\footnotesize {\bf A}, {\bf B}: The distribution of temporal covariances
    calculated in 100kb genomic windows from the \textcite{Barghi2019-qy}
    study, plotted alongside an empirical neutral null distribution created by
    recalculating the windowed covariances on a 1,000 sign permutations of
    allele frequency changes within tiles. The histogram bin number is 88,
    chosen by cross validation (Supplementary Materials
    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure {\bf A}, windowed
    covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by $k=2 \times
    10$ generations and in subfigure {\bf A} the covariances are separated by $k=4
    \times 10$ generations; each $k$ is an off-diagonal from the variance
    diagonal of the temporal covariance matrix (see cartoon of upper-triangle
    of covariance matrix in subfigures {\bf A} and {\bf B}, where the first diagonal is the
    variance, and the dark gray indicates which off-diagonal of the covariance
    matrix is plotted in the histograms). {\bf C}: The lower and upper tail
    probabilities of the observed windowed covariances, at 20\% and 80\%
    quintiles of the empirical neutral null distribution, for varying time
    between allele frequency changes (i.e. which off-diagonal $k$). The
  confidence intervals are  $95\%$ block-bootstrap confidence interval, and the
light gray dashed line indicates the 20\% tail probability expected under the
neutral null.}
    \end{center}
    \label{fig:figure-3} 
\end{figure}

@@

Finally, we observed that in the longest study  we analyzed
\parencite{Barghi2019-qy}, some genome-wide temporal covariances become
negative at future timepoints (see the first two rows in Figure
\ref{fig:figure-1} A). This shows that alleles that were on average going up
initially are later going down in frequency, i.e. that the average direction of
selection experienced by alleles has flipped. This must reflect either a change
in the environment or the genetic background, due to changing epistatic
relationships among alleles or  recombination breaking up selective alleles.
Such reversals of selective dynamics could be occurring at other timepoints but
the signal of a change in the direction of selection at particular loci may be
washed out when we calculate our genome-wide average temporal covariances.  To
address this limitation, we calculated the distribution of the temporal
covariances over 100kb genomic windows (Figure \ref{fig:figure-3}, pooling
across all replicates; see Supplementary Figure
\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
covariance of each tile will be noisy, due to sampling and genetic drift, and
the neutral distribution of the covariance is complicated due to linkage
disequilibria (which can occur over long physical distances in E\&R and
selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). To address
this, we have developed a permutation-based procedure that constructs a null
distribution by randomly flipping the signs of the  allele frequency changes
per-genomic window. This destroys the systematic covariances created by linked
selection and creates a sampling distribution of the covariances
spuriously created by neutral genetic drift while preserving the complex
dependencies between adjacent loci created by linkage disequilibrium (see
Supplementary Material Section \ref{supp:empirical-null} for more details on
this approach). This empirical neutral null distribution is conservative in
the sense that the variances of the covariances are wider than expected under
drift alone as they include the effect of selection on the allele frequency
change within a time-interval, just not between time-intervals. We see (Figure
\ref{fig:figure-3} A and B) that windowed temporal covariances between close
timepoints are skewed positive (a heavy right tail), while between more distant
timepoints these windowed temporal covariances tend to shift to become more
negative (a heavy left tail).  We quantified the degree to which the left and
right tails are inflated compared to the null distribution as a function of
time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
finding is also robust to sign-permuting allele frequency changes on a
chromosome-level, the longest extent that gametic linkage disequilibria can
extend (Supplementary Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a
striking pattern that the windowed covariances not only decay towards zero, but
in fact become negative through time, consistent with many regions in the
genome having had a reversed fitness effect at later timepoints.

@@

One limitation of our current work here is that none of the studies we
reanalyzed had linkage disequilibria data for the evolved populations. Our
theory of temporal autocovariance tells us that the temporal autocovariances a
neutral site experiences is determined by the product of the expected linkage
disequilibrium and additive fitness variation that persists through the
generations \parencite{Buffalo2019-io}. This leads to a clear prediction:
regions of higher linkage disequilibrium and lower recombination should have
greater temporal autocovariance than regions with lower LD and higher
recombination. Unfortunately, we lack a high-resolution recombination map for
\emph{D. simulans} and while there are LD data for \emph{D. simulans}
\parencite{Signor2018-wg} we did not find a relationship between temporal
covariance and LD.

@@

We believe this is driven by the idiosyncratic nature of LD in
evolve-and-resequence populations \parencite{Nuzhdin2013-gf,Kelly2019-dc}, and
that we might find such a relationship if LD data from the evolved population
were available. Unfortunately, sequencing multiple timepoints is expensive and
the low-coverage and/or pooled-sequencing approaches common in these studies
prohibit estimating linkage disequilibrium. Future studies complete with LD
data and recombination maps would allow one to disentangle the influence of
closely linked sites from more distant sites in causing temporal
autocovariance, and allow us better understand sites in high recombination
regions are free from the effects of linked selection. Furthermore, such
additional data would allow for localizing the effects of temporal covariance,
which would allow us to estimate not just the genome-wide fraction of variation
in frequency change ($G(t)$), but also the fraction of genome impacted by
linked selection, allowing us to synthesize the temporal approach with
single-timepoint studies like that of \textcite{Elyashiv2016-vt}.

@@

@@

\begin{figure}[!ht]
  \centering
  \includegraphics[]{figures/barghi-cov-panels.pdf}

@@

\subsection{\textcite{Bergland2014-ij} Re-Analysis}

>>>>

  Adaptation can proceed in natural populations and selection experiments over
  very short timescales.  However, detecting these changes genome-wide is
  difficult, as rapid adaptation often proceeds from standing variation or is
  the result of selection on polygenic traits, both of which may leave faint
  genomic signals indistinguishable from a noisy background of genetic drift.
  One promising signal comes from the genome-wide covariance between allele
  frequency changes observable from temporal genomic data, e.g.
  evolve-and-resequence studies. These temporal covariances reflect how the
  change in neutral allele frequency at one timepoint is predictive of the
  changes at later timepoints when there is heritable fitness variation in the
  population, as neutral alleles can remain associated with selected alleles
  over time. Since genetic drift does not lead to temporal covariance, we can
  use these covariances to estimate what fraction of the variation in allele
  frequency change through time is driven by linked selection. Here, we
  reanalyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
  one artificial selection experiment in mice, to quantify the effects of
  linked selection over short timescales.  Additionally, taking advantage of
  selection on replicated populations, we can estimate the covariance in allele
  frequency between replicates due to convergent selection pressure.  We also
  identify signals of negative temporal covariance corresponding to reversals
  in the direction of selection for a reasonable proportion of loci over the time
  course of a selection experiment.  Overall, we find that in the three studies
  we analyzed, linked selection has a significant impact on short-term allele
  frequency dynamics that previously may have been attributed to genetic drift

@@

A long-standing problem in evolutionary genetics is quantifying the roles of
genetic drift and selection in shaping genome-wide allele frequency changes.
Selection can both directly and indirectly affect allele frequencies, with the
indirect effect coming from the action of selection on correlated loci
elsewhere in genome e.g. linked selection (\cite{Maynard_Smith1974-lc},
\cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
review).  Previous work on this question has mostly focused on teasing apart
the impacts of drift and selection on genome-wide diversity using population
samples from a single contemporary timepoint, often by modeling the correlation
between regional recombination rate, gene density, and diversity created in the
presence of linked selection \parencite{Begun1992-ey,Elyashiv2016-vt}. This
approach has shown linked selection has a major role in shaping patterns of
genome-wide diversity across the genomes of some of organisms
\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\cite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt}.
However, we lack an understanding of how genome-wide linked selection acts over
time.

There are numerous examples of rapid phenotypic adaptation
\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid
genomic evolution in asexual populations
\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
of fitness makes detecting the impact selection on genome-wide variation over
short timescales in sexual populations remarkably difficult. This is because
the effect of selection on a polygenic trait (such as fitness) is distributed
across loci in proportion to their effect sizes. This can lead to subtle allele
frequency shifts on standing variation that are difficult to distinguish from
background levels of genetic drift and sampling variance.  However,
increasingly genomic experimental evolution studies with multiple timepoints,
and in some cases multiple replicate populations, are being used to detect
selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate modes
of selection \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In
addition these temporal-genomic studies have begun in wild populations, some
with the goal of finding variants that exhibit frequency changes consistent
with fluctuating selection \parencite{Bergland2014-ij,Machado2018-cs}. \vb{In a
  previous paper, we proposed that one useful signal for understanding the
  genome-wide impact of polygenic linked selection detectable from temporal
genomic studies} is the temporal autocovariance in allele frequency changes
\parencite{Buffalo2019-io}.  These covariances are directly estimable from
temporal genomic data and are created when the loci that underly heritable
fitness variation perturb the frequencies of linked neutral alleles; in
contrast, when genetic drift acts alone in a closed population, these
covariances are zero in expectation. Mathematically, temporal covariances are
useful because it is natural to decompose the total variance in allele
frequency change across a set of time intervals into the variances and
covariances in allele frequency change among time intervals.  Furthermore,
biologically, these covariances reflect the extent to which allele frequency
changes in one generation predict changes in another due to a shared fitness
background and shared selection pressures.

Here, we provide the first empirical analyses to quantify the impact of linked
selection acting over short timescales (tens of generations) across two evolve
and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and one
artificial selection experiment \parencite{Castro2019-uk}. We repeatedly find a
signal of temporal covariance, consistent with linked selection acting to
significantly perturb genome-wide allele frequency changes across the genome in
a manner that other approaches would not be able differentiate from genetic
drift. We estimate the lower bound on the proportion of total variation in
allele frequency change caused by linked selection, and the correlation between
allele frequency changes between replicate populations caused by response to
convergent selection pressures. Overall, we demonstrate that linked selection
has a powerful role in shaping genome-wide allele frequency changes over very
short timescales.

@@

\begin{center}
  \footnotesize
  \begin{tabular}{l c c c c c c}
    Study & Species & Selection & Replicates & Pop. Size & Gens. & Timepoints \\ 
  \hline
    \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab & 3 & $\sim$1100 & 14 & 2 \\ \hline
    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab & 10 & $\sim$1000 & 60 & 7 \\ \hline
    \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 20 & 2 \\
 \hline
\end{tabular}
\end{center}

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature lab environment and
evolved for 60 generations, and sequenced every ten generations. Using the
seven timepoints and ten replicate populations, we estimated the genome-wide $6
\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
replicates. Each row of these matrices represent the temporal covariance
$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
change (in ten-generation intervals, denoted $\Delta_{_{10}} p_t$) in some
initial reference generation $s$ (the row of the matrix), and some later
timepoint $t$ (the column of the matrix). We corrected these matrices for
biases created due to sampling noise, and normalize the entries for
heterozygosity (see Supplementary Material Sections
\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction} for details on
the bias correction). The covariances are expected to be zero when only drift
is acting in a closed population, as only heritable variation for fitness can
create covariance between allele frequency changes \parencite{Buffalo2019-io}.
Averaging across the ten replicate temporal covariances matrices, we find
temporal covariances that are statistically significant (95\% block bootstraps
CIs do not contain zero), consistent with linked selection perturbing
genome-wide allele frequency changes over very short time periods. The
covariances between all adjacent time intervals are positive and then decay
towards zero as we look at more distant time intervals, as expected when
directional selection affects linked variants' frequency trajectories until
ultimately linkage disequilibrium and the additive genetic variance for fitness
associated with neutral alleles decays \parencite{Buffalo2019-io}. The temporal
covariances per replicate are noisier but this general pattern holds; see
Supplementary Figures \ref{suppfig:barghi-cov-panels}.
\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
time intervals are on average ten generations apart, and given the temporal
decay in covariance we see, the covariances on shorter time-scales (e.g. if
adjacent generations had been sequenced) may well be higher yet (see
Supplementary Material Section \ref{supp:barghi-covs} for more details).

@@

While the presence of positive temporal covariances is consistent with linked
selection affecting allele frequencies over time, this measure not easily
interpretable. Additionally, we can quantify the impact of linked selection on
allele frequency change as the ratio of total covariance in allele frequency
change to the total variance in allele frequency change. \vb{We denote the
change in allele frequency as $\Delta p_t = p_{t+1}-p_t$, where $p_t$ is the
allele frequency in generation $t$}. Since the total variation in allele
frequency change can be partitioned into variance and covariance components,
$\var(p_t - p_0) = \sum_{i} \var(\Delta p_i) + \sum_{i \ne j} \cov(\Delta
p_i, \Delta p_j)$, and the covariances are zero when drift acts alone, this is
a lower bound on how much of the variance in allele frequency change is caused
by linked selection \parencite{Buffalo2019-io}. We call this measure $G(t)$, defined as

\begin{align}
  G(t) = \frac{\sum_{i\ne j}^t \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
\end{align}
%
\vb{which estimates} the effect of linked selection between the initial
generation $0$ and some later generation $t$, which can be varied to see how
this quantity grows through time. As with the temporal covariances, the study
design of \textcite{Barghi2019-qy} leads our measure $G(t)$ to be even more
conservative, since the temporal covariances within each ten-generation block
between sequenced timepoints are not directly observable, and are not included
in the numerator of $G(t)$. Still, we find a \vb{remarkably strong signal.
Greater} than $20\%$ of total variation in allele frequency change over 60
generations is the result of linked selection.

\vb{Additionally, we looked for a signal of temporal autocovariance in
\textcite{Bergland2014-ij}, a study that collected \emph{Drosophila
melanogaster} through Spring-Fall seasons across three years. With a pattern of
genome-wide fluctuating selection, we might expect a pattern of positive
covariances between similar seasonal changes, and negative covariances between
dissimilar seasonal changes. However, we find no such signal; we discuss this
in more depth in Supplementary Materials Section \ref{bergland-reanalysis}.}

The replicate design of \textcite{Barghi2019-qy} allows us to quantify another
covariance: the covariance in allele frequency change between replicate
populations experiencing convergent selection pressures. These
between-replicate covariances are created in the same way as temporal
covariances: neutral alleles linked to a particular fitness background
experience are expected to have allele frequency changes in the same direction
if the selection pressures are similar. Intuitively, where temporal covariances
reflect that neutral alleles associated with heritable fitness backgrounds are
predictive of frequency changes between generations, replicate covariances
reflect that heritable fitness backgrounds common to each replicate predict
(under the same selection pressures) frequency changes between replicates. We
measure this through a statistic similar to a correlation, which we call the
convergent correlation: the ratio of average between-replicate covariance
across all pairs to the average standard deviation across all pairs of
replicates,

@@

A benefit of between-replicate covariances is that unlike temporal covariances,
these can be calculated with only two sequenced timepoints and a replicated
study design. This allowed us to assess the impact of linked selection in
driving convergent patterns of allele frequency change across replicate
populations in two other studies. First, we reanalyzed the selection experiment
of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
environment. Since each replicate was exposed to the same selection pressure
and share linkage disequilibria common to the original natural founding
population, we expected each of the three replicate populations to have
positive between-replicate covariances. We find all three pairwise
between-replicate covariances are positive and statistically significant (95\%
CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
convergent correlation coefficient across these replicates as 0.36 ($95\%$
block-bootstrap confidence interval $[0.31, 0.40]$).

%\gc{TODO GIVE conversion \% var due to linked selection. }
Second, we reanalyzed the Longshanks selection experiment, which selected for
longer tibiae length relative to body size in mice, leading to a response to
selection of about 5 standard deviations over the course of twenty generations
\parencite{Castro2019-uk}. This study includes two independent selection lines,
Longshanks 1 and 2 (LS1 and LS2), and an unselected control line (Ctrl).
Consequently, this selection experiment offers a useful control to test our
between-replicate covariances: we expect to see positive between-replicate
covariance in the comparison between the two Longshanks selection lines, but
not between the two pairwise comparisons between the control line and the two
Longshanks lines. We find that this is the case (gray confidence intervals in
Figure \ref{fig:figure-2} C), with the two Longshanks comparisons to the
control line not being significantly different from zero, while the comparison
between the two Longshanks line is statistically significantly different from
zero (CIs [0.013, 0.037]).

@@

\begin{figure}
  \centering
  \includegraphics[]{figures/figure-3.pdf}

  \caption{\footnotesize {\bf A}, {\bf B}: The distribution of temporal covariances
    calculated in 100kb genomic windows from the \textcite{Barghi2019-qy}
    study, plotted alongside an empirical neutral null distribution created by
    recalculating the windowed covariances on a 1,000 sign permutations of
    allele frequency changes within tiles. The histogram bin number is 88,
    chosen by cross validation (Supplementary Materials
    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure {\bf A}, windowed
    covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by $k=2 \times
    10$ generations and in subfigure {\bf A} the covariances are separated by $k=4
    \times 10$ generations; each $k$ is an off-diagonal from the variance
    diagonal of the temporal covariance matrix (see cartoon of upper-triangle
    of covariance matrix in subfigures {\bf A} and {\bf B}, where the first diagonal is the
    variance, and the dark gray indicates which off-diagonal of the covariance
    matrix is plotted in the histograms). {\bf C}: The lower and upper tail
    probabilities of the observed windowed covariances, at 20\% and 80\%
    quintiles of the empirical neutral null distribution, for varying time
    between allele frequency changes (i.e. which off-diagonal $k$). The
  confidence intervals are  $95\%$ block-bootstrap confidence intervals, and the
light gray dashed line indicates the 20\% tail probability expected under the
neutral null.}
    \label{fig:figure-3} 
\end{figure}

@@

Finally, we observed that in the longest study  we analyzed
\parencite{Barghi2019-qy}, some genome-wide temporal covariances become
negative at future timepoints (see the first two rows in Figure
\ref{fig:figure-1} A). This shows that alleles that were on average going up
initially are later going down in frequency, i.e. that the average direction of
selection experienced by alleles has flipped. This must reflect either a change
in the environment or the genetic background, due to epistatic relationships
among alleles \vb{altered by frequency changes} or recombination breaking up
selective alleles.  Such reversals of selective dynamics could be occurring at
other timepoints but the signal of a change in the direction of selection at
particular loci may be washed out when we calculate our genome-wide average
temporal covariances.  To address this limitation, we calculated the
distribution of the temporal covariances over 100kb genomic windows (Figure
\ref{fig:figure-3}, pooling across all replicates; see Supplementary Figure
\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
covariance of each tile will be noisy, due to sampling and genetic drift, and
the neutral distribution of the covariance is complicated due to linkage
disequilibria (which can occur over long physical distances in E\&R and
selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). To address
this, we have developed a permutation-based procedure that constructs a null
distribution by randomly flipping the signs of the  allele frequency changes
per-genomic window. This destroys the systematic covariances created by linked
selection and creates a sampling distribution of the covariances spuriously
created by neutral genetic drift while preserving the complex dependencies
between adjacent loci created by linkage disequilibrium (see Supplementary
Material Section \ref{supp:empirical-null} for more details on this approach).
This empirical neutral null distribution is conservative in the sense that the
variances of the covariances are wider than expected under drift alone as they
include the effect of selection on the allele frequency change within a
time-interval, just not between time-intervals. We see (Figure
\ref{fig:figure-3} A and B) that windowed temporal covariances between close
timepoints are skewed positive (a heavy right tail), while between more distant
timepoints these windowed temporal covariances tend to shift to become more
negative (a heavy left tail).  We quantified the degree to which the left and
right tails are inflated compared to the null distribution as a function of
time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
finding is also robust to sign-permuting allele frequency changes on a
chromosome-level, the longest extent that gametic linkage disequilibria can
extend (Supplementary Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a
striking pattern that the windowed covariances not only decay towards zero, but
in fact become negative through time, consistent with many regions in the
genome having had a reversed fitness effect at later timepoints.

@@

One limitation \vb{of these analyses is} that none of the studies we reanalyzed
estimated linkage disequilibria data for the evolved populations. Our theory of
temporal autocovariance tells us that the temporal autocovariances a neutral
site experiences is determined by the product of the expected linkage
disequilibrium and additive fitness variation that persists through the
generations \parencite{Buffalo2019-io}. This leads to a clear prediction:
regions of higher linkage disequilibrium and lower recombination should have
greater temporal autocovariance than regions with lower LD and higher
recombination. Unfortunately, we lack a high-resolution recombination map for
\emph{D. simulans} and while there are LD data for \emph{D. simulans}
\parencite{Signor2018-wg} we did not find a relationship between temporal
covariance and LD.

@@

We believe this is driven by the idiosyncratic nature of LD in
evolve-and-resequence populations \parencite{Nuzhdin2013-gf,Kelly2019-dc}, and
that we might find such a relationship if LD data from the evolved population
were available. Unfortunately, sequencing multiple timepoints is expensive and
the low-coverage and/or pooled-sequencing approaches common in these studies
prohibit estimating linkage disequilibrium. Future studies complete with LD
data and recombination maps would allow one to disentangle the influence of
closely linked sites from more distant sites in causing temporal
autocovariance, and allow us to better understand whether sites in high
recombination regions are free from the effects of linked selection.
Furthermore, such additional data would allow for localizing the effects of
temporal covariance, which would allow us to estimate not just the genome-wide
fraction of variation in frequency change ($G(t)$), but also the fraction of
the genome impacted by linked selection, allowing us to synthesize the temporal
approach with single-timepoint studies like that of \textcite{Elyashiv2016-vt}.

@@

\printbibliography

@@

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{figures/barghi-cov-panels.pdf}

@@

\subsection{\textcite{Bergland2014-ij} Re-Analysis}
\label{bergland-reanalysis}
